Modular Curves

### 2020-01-22 paper.tex
introduced moduli spaces

This $\tau$ however may not be unique. Suppose that for some $\tau' \in \mathcal{H} $ the torus $\C / \Lambda_{\tau'}$ is also isomorphic to $\C / \Lambda$. In that cause $\C / \Lambda_{\tau'}$ is also isomorphic to $\C / \Lambda_{\tau}$. 
Then for some $m$ it must be the case that $m \Lambda_{\tau'} = m\tau' \Z \oplus m \Z = \Lambda_\tau$. 
By \cref{lem:equality_of_lattices} this implies that there is some $\gamma = \begin{pmatrix} a & b \\ c& d \end{pmatrix}  \in \SL_2\Z$ \[
\begin{pmatrix} m\tau' \\ m \end{pmatrix}  = \begin{pmatrix} a & b \\ c& d \end{pmatrix}  \begin{pmatrix} \tau \\ 1 \end{pmatrix} 
.\] 
So \[
\tau' = \frac{m \tau'}{m} = \frac{a \tau + b}{c \tau + d} = \gamma\cdot \tau
.\] 
Hence we see that $\tau$ and $\tau'$ lie in the same orbit of the action of $\SL_2\Z$. 
Converesely $\C / \Lambda_\tau$ and $\C / \Lambda_{\gamma\cdot \tau}$ will always be isomorhic.
Hence we conclude that there is a bijctice correspondence of isomorphism classes of complex tori and orbits of $\SL_2\Z$ acting on $\mathcal{H} $.\[
\begin{tikzcd}
	\left\{ \substack{\text{isormorphism classes} \\ \text{of complex tori}} \right\} \arrow[r, leftrightarrow] & 
	\mathcal{H} / \SL_2\ 
\end{tikzcd}
.\] 
\subsection{Isogenies}
If we only consider the maps where the contsant term $b$ is $0$ then these maps are clearly group morphisms between $\C / \Lambda$ and  $\C / \Lambda'$. 
As a result, the a holomorphic group morphism between these two tori exists if and only if $m\Lambda = \Lambda'$ for some $m \in \C$. 
A nonzero holomorphic group morphism between complex tori is called an \emph{isogeny}.

@@

From now on we will often say complex elliptic curve, instead of complex torus to highlight their interpretation as elliptic curves.

@@

\section{Appendices}
\begin{theorem}
	The modular group $\SL_2(\Z)$ is generated by \[ 	
	\begin{pmatrix}  1 & 1 \\ 0 & 1 \end{pmatrix} 
	\text{ and }
	\begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix} 
.\] 
\end{theorem}
\begin{proof}
	Let $\mathcal{A} $ be the group generated by the two matrices. 
	Suppose that $\SL_2(\Z) \setminus \mathcal{A} $ is not empty.
	Then we can choose a matrix \[
		A =\begin{pmatrix} a & b \\ c& d\end{pmatrix} \in \SL_2\Z \setminus \mathcal{A} 
	\] 
	with mimimal 1-norm, i.e. $|a| + |b| + |c| + |d|$ is minimal.
	We know that 
\end{proof}
\begin{exercise}[exercise 1.4.1 in \cite{diamondFirstCourseModular2005a}] \label{ex:1.4.1}
	Let $\Lambda = \omega_1 \Z + \omega_2 \Z$ be a lattice,  $f$ be a non-constant elliptic function on $\C / \Lambda$ and $P = \{x \omega_1 + y\omega_2 \;|\: x, y \in [0,1]\} $ be a fundamental parallellogram of $\Lambda$.
	 \begin{enumerate}
		\item Proof that the sum of residues on $f$ is 0. 
		\item Proof that $f$ has as many poles as zeroes (counting multiplicity). 
			Conclude that $f$ takes any particular value exact  $n$ times (counting multiplicity) for some fixed $n$. Show that in particular  $\wp_\Lambda$ takes any value twice.
		\item Show that 
			\[
				\sum_{x \in E} \nu_x(f) x \in \Lambda
			,\]
			where $\nu_x(f)$ is the order of $f$ at  $x$. 
	\end{enumerate}
\end{exercise}
\begin{proof}
	Let $\partial P$ be the counter clockwise oriented boundary of $P$. Because $f$ has a finite number of poles and zeros we can translate this boundary by some $t$ such that $\partial P + t$ does not contain any poles or zeroes.
\begin{figure}[h]
	\incfig{ex141}	
\end{figure}
\begin{enumerate}
	\item First note that boundary integral can be split up by integrating over the edges of $P + t$ separately\[
			\int_{\partial P+ t} f(z) \mathrm d z = \left(\int_t^{\omega_2 + t} + \int_{\omega_2+t}^{\omega_1 + \omega_2 + t} + \int_{\omega_1 + \omega_2 + t} ^{\omega_1 + t} + \int_{\omega_1}^{t}\right) f(z) \mathrm dz
	.\] 
	Note that the integrals corresponding with the two blue edges in the picture will cancel by the $\Lambda$ periodicity of $f$. Similarly the two red edges cancel. 
	Hence \[
		\int_{\partial P + t} f(z) \mathrm d z = 0
	.\] 
	By the residue theorem we find that the sum of residues on $f$ is zero.

@@

The Weierstrass function $\wp_{\Lambda}$ has only one pole of order 2. Hence in this case  $n = 2$.
\item 
	This follows directly from the generalised argument principle applied to both \[
		\int_{\partial P _+ t} z \frac{f(z)}{f'(x)} \mathrm d z \text{ and } \int_{\partial P + t} \frac{f(z)}{f'(z)} \mathrm d z
	.\] 
	The second integral is $0$ as discussed above. 
	The first integral adds only one pole at $z = 0$. Hence  \[
		\int_{\partial P + t} z \frac{f(z)}{f'(z)}\mathrm d z = \sum_
	.\]\todo{What the heck?}
\end{enumerate}
\end{proof}
\printbibliography
\end{document}

>>>>

This $\tau$ however may not be unique. Suppose that for some $\tau' \in \mathcal{H} $ the torus $\C / \Lambda_{\tau'}$ is also isomorphic to $\C / \Lambda$. In that cause $\C / \Lambda_{\tau'}$ is also isomorphic to $\C / \Lambda_{\tau}$. 
Then for some $m$ it must be the case that $m \Lambda_{\tau'} = m\tau' \Z \oplus m \Z = \Lambda_\tau$. 
By \cref{lem:equality_of_lattices} this implies that there is some $\gamma = \begin{pmatrix} a & b \\ c& d \end{pmatrix}  \in \SL_2\Z$ \[
\begin{pmatrix} m\tau' \\ m \end{pmatrix}  = \begin{pmatrix} a & b \\ c& d \end{pmatrix}  \begin{pmatrix} \tau \\ 1 \end{pmatrix} 
.\] 
So \[
\tau' = \frac{m \tau'}{m} = \frac{a \tau + b}{c \tau + d} = \gamma\cdot \tau
.\] 
Hence we see that $\tau$ and $\tau'$ lie in the same orbit of the action of $\SL_2\Z$. 
Converesely $\C / \Lambda_\tau$ and $\C / \Lambda_{\gamma\cdot \tau}$ will always be isomorhic.
Hence we conclude that there is a bijectice correspondence of isomorphism classes of complex tori and orbits of $\SL_2\Z$ acting on $\mathcal{H} $.\[
\begin{tikzcd}
	\left\{ \substack{\text{isormorphism classes} \\ \text{of complex tori}} \right\} \arrow[r, leftrightarrow] & 
	\mathcal{H} / \SL_2\Z\ 
\end{tikzcd}
.\] 
\subsection{Isogenies}
If we only consider the maps where the contsant term $b$ is $0$ then these maps are clearly group morphisms between $\C / \Lambda$ and  $\C / \Lambda'$. 
As a result, the a holomorphic group morphism between these two tori exists if and only if $m\Lambda = \Lambda'$ for some $m \in \C$. 
A nonzero holomorphic group morphism between complex tori is called an \emph{isogeny}.

@@

From now on we will often say complex elliptic curve, instead of complex torus to highlight their interpretation as elliptic curves. Further, we will often denote complex elliptic curves with a simple symbold (often $E$) instead of always writing them as the quotient  $\C / \Lambda$.

@@

In \cref{sec:modular_forms_and_complex_tori} we have seen that there is a bijective correspondence between equivaleny classes of complex elliptic curves and points in $Y (\SL_2\Z)$. 
One may wonder whether there there is a similar result for modular curves $Y_0(N)$ and $Y_1(N)$.
It turns out there is. 
Instead of restricting the isomorphisms between the complex elliptic curves, we will actually give the elliptic curves extra information, resulting in \emph{enhanced elliptic curves}.

\subsection{Enhanced Elliptic Curves and Moduli Spaces} \label{sec:enhanced_elliptic_curves_and_moduli_spaces}
There are three common ways of `adding information' to elliptic curves, creating \emph{enhanced elliptic curves}. Moduli Spaces are equivalence classes of enhanced elliptic curves.  
\begin{definition} \hspace{\linewidth}
	\begin{enumerate}
		\item An \emph{enhanced elliptic curve for $\Gamma_0(N)$} is an pair $(E, C)$ where  $E$ is a complex elliptic curve and $C$ is an cyclic subgroup of order $N$.
		Two of these curves $(E_1, C_1), (E_2, C_2)$ are equivalent if and only if there is an isomorphism from $E_1$ to $E_2$ that restricts to an isomorphism from $C_1$ to $C_2$.
		The associated \emph{moduli space} is \[
			S_0(N) = \{ \text{enhanced elliptic curves for } \Gamma_0(N)\} / \sim
		.\] 
	\item An \emph{enhanced elliptic curve for $\Gamma_1(N)$} is an pair $(E, Q)$ where  $E$ is a complex elliptic curve and $Q$ is point in $E$ of order $N$.
		Two of these curves $(E_1, Q_1), (E_2, Q_2)$ are equivalent if and only if there is an isomorphism from $E_1$ to $E_2$ that maps $Q_1$ to $Q_2$.
		The associated \emph{moduli space} is \[
			S_1(N) = \{ \text{enhanced elliptic curves for } \Gamma_0(N)\} / \sim
		.\] 
	\item An \emph{enhanced elliptic curve for $\Gamma(N)$} is an tripple $(E, P, Q)$ where  $E$ is a complex elliptic curve and $P, Q$ are points in the $N$-torsion subgroup $E[N]$ such  that the weil pairing  $e_N(P, Q) = 1$.
		Two of these curves $(E_1, P_1, Q_1), (E_2, P_2, Q_2)$ are equivalent if and only if there is an isomorphism from $E_1$ to $E_2$ that maps $P_1$ to $P_2$ and $Q_1$ to $Q_2$
		The associated \emph{moduli space} is \[
			S(N) = \{ \text{enhanced elliptic curves for } \Gamma_0(N)\} / \sim
		.\] 
	\end{enumerate}	
\end{definition}
\Cref{fig:enhanced_elliptic_curves} gives a visualisation of these enhanced elliptic curves.
\begin{figure}[h]
	\centering
	\incfig{enhanced_elliptic_curves}
	\caption{Enhanced elliptic curves}
	\label{fig:enhanced_elliptic_curves}
\end{figure}
The following theorem gives a generalisation of the conclusion of \cref{sec:modular_curves_and_moduli_spaces}, showing that there are bijections in the modular curves $Y_0(N), Y_1(N), Y(N)$ and the moduli spaces $S_0(N), S_1(N), S(N)$, as well as providing more insight in what the moduli spaces look like. 
\begin{theorem}
	\hspace{\linewidth}
	\begin{enumerate}
		\item Every enhanced elliptic curve for $\Gamma_0(N)$ is equivalent to $(E_\tau, \left<1 / N + \Lambda_\tau \right>$ 
	\end{enumerate}
\end{theorem}

\section{Appendices}
\begin{theorem}
	The modular group $\SL_2(\Z)$ is generated by \[ 	
	\begin{pmatrix}  1 & 1 \\ 0 & 1 \end{pmatrix} 
	\text{ and }
	\begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix} 
.\] 
\end{theorem}
\begin{proof}
	Let $\mathcal{A} $ be the group generated by the two matrices. 
	Suppose that $\SL_2(\Z) \setminus \mathcal{A} $ is not empty.
	Then we can choose a matrix \[
		A =\begin{pmatrix} a & b \\ c& d\end{pmatrix} \in \SL_2\Z \setminus \mathcal{A} 
	\] 
	with mimimal 1-norm, i.e. $|a| + |b| + |c| + |d|$ is minimal.
	We know that 
\end{proof}
\begin{exercise}[exercise 1.4.1 in \cite{diamondFirstCourseModular2005a}] \label{ex:1.4.1}
	Let $\Lambda = \omega_1 \Z + \omega_2 \Z$ be a lattice,  $f$ be a non-constant elliptic function on $\C / \Lambda$ and $P = \{x \omega_1 + y\omega_2 \;|\: x, y \in [0,1]\} $ be a fundamental parallellogram of $\Lambda$.
	 \begin{enumerate}
		\item Proof that the sum of residues on $f$ is 0. 
		\item Proof that $f$ has as many poles as zeroes (counting multiplicity). 
			Conclude that $f$ takes any particular value exact  $n$ times (counting multiplicity) for some fixed $n$. Show that in particular  $\wp_\Lambda$ takes any value twice.
		\item Show that 
			\[
				\sum_{x \in \C / \Lambda} \nu_x(f) x \in \Lambda
			,\]
			where $\nu_x(f)$ is the order of $f$ at  $x$. 
	\end{enumerate}
\end{exercise}
\begin{proof}
	Let $\partial P$ be the counter clockwise oriented boundary of $P$. Because $f$ has a finite number of poles and zeros we can translate this boundary by some $t$ such that $\partial P + t$ does not contain any poles or zeroes.
\begin{figure}[h]
	\incfig{ex141}	
\end{figure}
\begin{enumerate}
	\item First note that boundary integral can be split up by integrating over the edges of $P + t$ separately\[
			\int_{\partial P+ t} f(z) \mathrm d z = \left(\int_t^{\omega_2 + t} + \int_{\omega_2+t}^{\omega_1 + \omega_2 + t} + \int_{\omega_1 + \omega_2 + t} ^{\omega_1 + t} + \int_{\omega_1}^{t}\right) f(z) \mathrm dz
	.\] 
	Note that the integrals corresponding with the two blue edges in the picture will cancel by the $\Lambda$ periodicity of $f$. Similarly the two red edges cancel. 
	Hence \[
		\int_{\partial P + t} f(z) \mathrm d z = 0
	.\] 
	By the residue theorem we find that the sum of residues on $f$ is zero.

@@

The Weierstrass function $\wp_{\Lambda}$ has only one pole of order 2. Hence in this case  $n = 2$.
\item 
	We will show this by first calculating \[
		\frac{1}{2\pi i}\int_{\partial P + t} z \frac{f(z)}{f'(z)} \mathrm d z
	\]
	and showing that this value must be in $\Lambda$.
	Like in the first part of this exercise we will split up the boundary in its edges.
	\[	
		\int_{\partial P+ t} z \frac{f(z)}{f'(z)} \mathrm d z = \left(\int_t^{\omega_2 + t} + \int_{\omega_2+t}^{\omega_1 + \omega_2 + t} + \int_{\omega_1 + \omega_2 + t} ^{\omega_1 + t} + \int_{\omega_1}^{t}\right) z \frac{f(z)}{f'(z)} \mathrm dz
	.\]
	By $\Lambda$-periodicity we know that the factor $f(z) / f'(z)$ is the same on opposing edges. Hence we can write \begin{align*}
		\int_{\partial P + t} z \frac{f(z)}{f'(z)} \mathrm d z &= \int_t^{\omega_1 + t} (z + \omega_2 - z) \frac{f(z)}{f'(z)} \mathrm d z + \int_{t}^{\omega_2 + t} (z - z - \omega_1) \frac{f(z)}{f'(z)} \mathrm d z\\
								       &= \omega_2 \int_t^{\omega_1 + t} 1 \; \mathrm d \ln f(z) - \omega_1 \int_t^{\omega_2 +t} 1 \; \mathrm d \ln f(z) \\
	.\end{align*} 
	From this we see that the integrals must be integer multiples of  $2\pi i$. 
	Hence 
	\[
		\frac{1}{2\pi i } \int_{\partial P + t} z \frac{f(z)}{f'(z)} \mathrm d z \in \Lambda
	.\] 
	Now note that we can write $f(z) = \sum_{x \in \{\text{poles or zero}\} }(z - x)^{\nu_x(f)}g(z)$ for some some simple function $g$. 
	So \[
		\frac{zf'(z)}{f(z)} = \sum_{x \in \{\text{poles or zero}\} }\nu_x(f) \frac{z}{ z - x} + z \frac{g'(z)}{g(z)}
	.\] 
	Note that this function has poles with residue $x \nu_x(f)$ for every $x$ which is a pole or zero of  $f$. By the residue theorem  \[
		\frac{1}{2\pi i} \int_{\partial P + t} z \frac{f'(z)}{f(z)} \mathrm d z = \sum_{x \in \C / \Lambda} \nu_x(f)  x \in \Lambda 
	.\]  
\end{enumerate}
\end{proof}
\printbibliography
\end{document}
